---
title: "Week 1: Discussion"
subtitle: "Making maps with `tmap` and beyond"
author: "Courtney Lorey"
date: 2025-10-2
---

In this section, we’ll kick-off our mapmaking skills using the tmap package. Your task will be to make a map of Rapa Nui (Easter Island or Isla de Pascua). Rapa Nui is a volcanic island in Polynesia and territory of Chile. It is widely known for monumental statues called moai, but it is also a place of geoscientific inquiry. Several research programs about the island’s volcanic history, petrology, and human-environment interactions exist.

# Learning Objectives
- Read in spatial objects with st_read() and read_stars()
- Create map with single spatial object
- Create map with multiple spatial objects
- Use different types of tmap plotting formats (e.g. tm_polygons(), tm_fill(), tm_dots(), etc.)
- Include essential map elements (e.g. scale bar & north arrow or graticules)
- Create an interactive map
- Challenge: Reproduce map using ggplot2 instead of tmap


```{r}
# Load the necessary packages

library(here)
library(tidyverse)
library(sf)
library(stars)
library(tmap)
```

```{r}
# Load the data

ei_points <- st_read(here("Weekly_Discussions", "week1_Making_Maps", "data", "ei_points.gpkg")) |> 
  filter(type == "volcano")

ei_elev <- read_stars(here("Weekly_Discussions", "week1_Making_Maps", "data", "ei_elev.tif"))

ei_borders <- st_read(here("Weekly_Discussions", "week1_Making_Maps", "data", "ei_border.gpkg"))

ei_roads <- st_read(here("Weekly_Discussions", "week1_Making_Maps", "data", "ei_roads.gpkg"))
```

```{r}
# Create a map of the outline Rapa Nui using tm_shape() + tm_borders()

tm_shape(ei_borders) + 
    tm_borders() + 
    tm_title("Easter Island")
```

```{r}
# Create a multiple-object map with tmap
# …denote the island’s borders and continuous elevation
# …denote the island’s volcanoes and roads
# …play with the color palette and essential map elements

tm_shape(ei_elev) + 
    tm_graticules(col = "lightgray") +
    tm_raster(col.scale = tm_scale_continuous(values = "-rd_yl_gn"),
              col.legend = tm_legend(
                           title = "Elevation (m asl)", 
                           orientation = "portrait", 
                           position = tm_pos_out("right")
                           )
                           ) + 
tm_shape(ei_borders) +
    tm_borders() + 
tm_shape(ei_roads) +
    tm_lines() + 
tm_shape(ei_points) + 
    tm_symbols(shape = 24, 
               size = "elevation", 
               size.legend = tm_legend(title = "Volcanoes (m asl)", orientation = "portrait", position = tm_pos_out("right"))) +
tm_compass(position = c("right", "top")) +
tm_scalebar() + 
tm_title("Easter Island")

```

```{r}
# Make it interactive! 

tmap_mode("view")

tm_shape(ei_elev) + 
    tm_raster(col.scale = tm_scale_continuous(values = "-rd_yl_gn"),
              col.legend = tm_legend(
                           title = "Elevation (m asl)", 
                           orientation = "portrait", 
                           position = tm_pos_out("right")
                           )
                           ) + 
tm_shape(ei_borders) +
    tm_borders() + 
tm_shape(ei_roads) +
    tm_lines() + 
tm_shape(ei_points) + 
    tm_symbols(shape = 24, 
               fill = "black",
               size = "elevation", 
               size.legend = tm_legend(title = "Volcanoes (m asl)", orientation = "portrait", position = tm_pos_out("right"))) +
tm_compass() +
tm_scalebar() + 
tm_title("Easter Island")

```